---
title: "Tutorial: Use Cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Use Cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}f
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. Installing `LHTpicker` package
For this vigenette to work `LHTpicker` must be installed and exist in R's default libpath.
```{r setup}

# devtools::install_github("https://github.com/d2gex/LHTPicker", dep=TRUE)

```

# 2. Fetch wanted LHTs from FishLife
There are two use cases for using FishLife to obtain missing Life History Traits (LHTs):
i) You want to evaluate a species for which no biological parameters are available, meaning you need to retrieve FishLife-predicted LHTs for the desired taxon.
ii) You have some biological parameters but not all, and you use FishLife to generate a new set of parameters that integrates and updates your existing data.

This section deals with *case i*.

## 2.1 Inputted taxa details

A csv file with the taxon names and the **wanted** LHTs you are seeking is required.The csv shows as columns those LHTs
you are after, as shown below.
```{r provide_fishlife_with_wanted_lhts}
taxon_lhts_to_fetch <- readRDS("data/wanted_taxon_details.rds")
head(taxon_lhts_to_fetch)
```
The wrapper uses a list to map column names in your csv to the actual names used by FishLife. Should some predicted
variable names change in FishLife in the future, you may need to tweak the list. See that such a list is a member of an
`R6` object and it is called `fishlife_context`. If nothing has changed in FishLife, unless you need to add more wanted parameters,
nothing else is required to do.
```{r lht_mapping_list}
LHTpicker::fishlife_context$lht_names
```
Likewise, to back-transform the obtained predicted variables `LHTpicker` provides a function that map such predicted variables
with the required function to convert it. As previously mentioned, you may not need to touch this list either.

```{r back_transformed_function_list}
LHTpicker::fishlife_context$backtransform_function_list
```

## 2.2 Fetching predicted LHTs
Please refer to the [reference](https://d2gex.github.io/LHTpicker/reference/PredictedLHTPicker.html) for further details.

```{r fetch_predicted_lhts}
p_lht_picker <- LHTpicker::PredictedLHTPicker$new(FishLife::FishBase_and_Morphometrics,
                                                  LHTpicker::fishlife_context$lht_names,
                                                  LHTpicker::fishlife_context$backtransform_function_list,
                                                  taxon_lhts_to_fetch)
predicted_lht_df <- p_lht_picker$pick_and_backtransform()
head(predicted_lht_df)
```
In case FishLife cannot provide details about a species for whatever reason, `LHTpicker` will return a column of NA
values for the rows holding details about the missing species, as shown below.
```{r fetch_predicted_lhts_non_existing_speices}
non_existant_taxon_lhts_to_fetch <- dplyr::mutate(taxon_lhts_to_fetch, taxon = dplyr::case_when(
    taxon == "Trisopterus luscus" ~ "IDoNoExist",
    .default = taxon
))
p_lht_picker <- LHTpicker::PredictedLHTPicker$new(FishLife::FishBase_and_Morphometrics,
                                                  LHTpicker::fishlife_context$lht_names,
                                                  LHTpicker::fishlife_context$backtransform_function_list,
                                                  non_existant_taxon_lhts_to_fetch)
predicted_lht_df <- p_lht_picker$pick_and_backtransform()
head(predicted_lht_df)
```

# 3 Fetch updated LHTs given new data

This section deals with *case ii*, that is, with fetching updated LHTs from FishLife.

## 3.1 Inputted predicting taxa details
A csv file with the taxon names you are after and the **wanted-to-update** LHTs is required as shown below. Let's suppose
we have most of the LHTs however we do not know the natural mortality(M) nor the age of maturity.
```{r}
taxon_lhts_to_update <- readRDS("data/wanted_update_taxon_details.rds")
head(taxon_lhts_to_update)
```
## 3.2 Fetching updated LHTs
Unless the previous function, we need to provide two extra parameters: a prefix that `LHTpicker` will use to concatenate
to the named of updated columns in the csv and function transformation list, similar to the back-transformation one, to transform
the provided LHTs into Fishlife's expected domain. Once again, unless you need other variables from FishLife, you do not
need to modify neither of the related transformation and back-transformation lists.
```{r}

u_lht_picker <- LHTpicker::UpdatedLHTPicker$new(
  FishLife::FishBase_and_Morphometrics,
  taxon_lhts_to_update,
  LHTpicker::fishlife_context$updated_prefix,
  LHTpicker::fishlife_context$transform_function_list,
  LHTpicker::fishlife_context$backtransform_function_list,
  LHTpicker::fishlife_context$lht_names
)
updated_lht_df <- u_lht_picker$pick_and_backtransform()
head(updated_lht_df)
```
See how each updated column name from FishLife has a prefix attached to it and how `updated_M` and `Amat` have now
been provided. In order to use the these LHTs for your assessment you need to use the whole returned set, to avoid breaking
the covariance matrix among parameters (Thorson et al., 2017)
