---
title: "Tutorial: Use Cases"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: Use Cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# 1. Install `LHTpicker`

To run this vignette, install `LHTpicker` in one of R's library paths.

```{r setup}
# devtools::install_github("d2gex/LHTpicker", dependencies = TRUE)
```

# 2. Retrieve LHTs from FishLife

LHTpicker supports two workflows:

1. Retrieve FishLife predictions when trait values are unavailable.
2. Update FishLife predictions using a partial set of supplied trait values.

This section covers the first workflow.

## 2.1 Input taxa and requested traits

Provide a data frame with a `taxon` column and one column for each requested LHT. In practice, this can be read from CSV. The requested-trait columns are `NA` in this example.
```{r provide_fishlife_with_wanted_lhts}
taxon_lhts_to_fetch <- readRDS("data/wanted_taxon_details.rds")
head(taxon_lhts_to_fetch)
```
`fishlife_context$lht_names` maps LHTpicker column names to FishLife field names. You only need to change it when requesting a trait outside the default set or when FishLife changes a field name.
```{r lht_mapping_list}
LHTpicker::fishlife_context$lht_names
```
`backtransform_function_list` maps each FishLife field to the function that converts it back to the user-facing scale. The default list normally requires no changes.

```{r back_transformed_function_list}
LHTpicker::fishlife_context$backtransform_function_list
```

## 2.2 Retrieve predicted LHTs

See the [reference documentation](https://d2gex.github.io/LHTpicker/reference/PredictedLHTPicker.html) for the full interface.

```{r fetch_predicted_lhts}
p_lht_picker <- LHTpicker::PredictedLHTPicker$new(FishLife::FishBase_and_Morphometrics,
                                                  LHTpicker::fishlife_context$lht_names,
                                                  LHTpicker::fishlife_context$backtransform_function_list,
                                                  taxon_lhts_to_fetch)
predicted_lht_df <- p_lht_picker$pick_and_backtransform()
head(predicted_lht_df)
```
If FishLife cannot match a taxon, LHTpicker preserves its input row and leaves the requested LHT values as `NA`.

```{r fetch_predicted_lhts_non_existing_species}
non_existent_taxon_lhts_to_fetch <- dplyr::mutate(taxon_lhts_to_fetch, taxon = dplyr::case_when(
    taxon == "Trisopterus luscus" ~ "IDoNoExist",
    .default = taxon
))
p_lht_picker <- LHTpicker::PredictedLHTPicker$new(FishLife::FishBase_and_Morphometrics,
                                                  LHTpicker::fishlife_context$lht_names,
                                                  LHTpicker::fishlife_context$backtransform_function_list,
                                                  non_existent_taxon_lhts_to_fetch)
predicted_lht_df <- p_lht_picker$pick_and_backtransform()
head(predicted_lht_df)
```

# 3. Update LHTs with supplied data

This section covers the second workflow: updating FishLife predictions with supplied data.

## 3.1 Input taxa and supplied traits

Provide a data frame with one taxon per row and the available LHTs in the remaining columns. In this example, natural mortality (`M`) and age at maturity (`Amat`) are missing.
```{r}
taxon_lhts_to_update <- readRDS("data/wanted_update_taxon_details.rds")
head(taxon_lhts_to_update)
```
## 3.2 Retrieve updated LHTs

Compared with the prediction workflow, this call also needs an `updated` prefix and a transformation list. LHTpicker prefixes each new LHT column with `updated_`. The transformation list converts supplied values to FishLife's internal scale; the back-transformation list converts the returned values to the user-facing scale. You normally do not need to change either list.
```{r}

u_lht_picker <- LHTpicker::UpdatedLHTPicker$new(
  FishLife::FishBase_and_Morphometrics,
  taxon_lhts_to_update,
  LHTpicker::fishlife_context$updated_prefix,
  LHTpicker::fishlife_context$transform_function_list,
  LHTpicker::fishlife_context$backtransform_function_list,
  LHTpicker::fishlife_context$lht_names
)
updated_lht_df <- u_lht_picker$pick_and_backtransform()
head(updated_lht_df)
```
Each new LHT column has the `updated_` prefix; for example, `updated_M` and `updated_Amat` are now available. For an assessment, use the complete updated trait set rather than mixing input and updated values, to preserve the covariance structure among parameters (Thorson et al., 2017).
